---
title: "Embedding Shiny in Quarto"
format: html
runtime: shiny_prerendered
---

# Introduction

This is an example of embedding a Shiny app within a Quarto document.

```{r setup}
library(shiny)
```

```{r}
#| context: server
function(input, output, session) {
  output$distPlot <- renderPlot({
    hist(rnorm(input$obs))
  })
}

```

```{r}
#| context: ui
fluidPage(
  titlePanel("Hello Shiny!"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("obs", "Number of observations:",
                  min = 1, max = 1000, value = 500)
    ),
    mainPanel(
      plotOutput("distPlot")
    )
  )
)

```


### Steps to Render the Document

1. **Open RStudio**: Open your Quarto project in RStudio.
2. **Switch to Source View**: Ensure you are in the Source View by clicking the "Source" tab at the top right of the editor window.
3. **Copy and Paste**: Copy the corrected content above and paste it into your `.qmd` file.
4. **Save the Document**: Save your Quarto document.
5. **Render the Document**: Render your document using the command:
   ```sh
   quarto render newtst.qmd